\par\break\null%
\vskip 0.1\vsize%
\section*{Preface}
\addcontentsline{toc}{chapter}{Preface}

This thesis marks the completion of the requirements for my Master's degree in Information Security at the University of Oslo, and the end of a total of five years of study at the same university. The topic of this thesis is discussed and developed together with my supervisor Håkon Jacobsen.


\subsubsection*{Acknowledgements}

I would like to express my sincere gratitude to my supervisor, Håkon Jacobsen. Thank you for your guidance and support over the past two years. Your genuine interest in the topic and curiosity have inspired and motivated me when I needed it most. I greatly appreciate your time and effort in guiding me through this project and all the discussions we have had over the past couple of years. You have given me a deeper understanding of security.

I would also like to thank my other supervisor, Nils Gruschka. I might not have developed an interest in security if it were not for you and the various courses you have taught over the past years. Thank you.

Next, I thank my family for continuously being interested in science and understanding how things work. Growing up with this attitude has given me the joy of being exposed to new ideas and a curiosity about how technology works. Thank you for your support throughout my life.

Last but not least, I would like to thank all my friends, especially those who can simply turn any bad day into a good one by being present. Thank you for the happy memories created over the past five years at the University of Oslo. You all know who you are. Thank you for being a source of joy and support throughout my time here. The past five years would not have been the same without you. Also, thanks to my friends Thanh Thao Le and Vetle Håland Bergstad for all the snacks, lunches, and support and for providing feedback on my thesis. Thank you all.

\vfill
\hfill \textsc{Embrik Hafskjold Thoresen}

\hfill \textit{Oslo, November 2024}
\vfill
